Polynomial phase estimation by phase unwrapping

### 2012-10-14 paper.tex
modified plots.  edits.

Polynomial phase signals arise in numerous fields including radar, sonar, geophysics, biology, and radio communication \cite{Angeby_estimating_2000}. In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar~\cite{Levanon_Radar_signals_2004}.  In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location \cite{Suga_1975_bats_echolocation, Moss_2005echolocation}.

A polynomial phase signal of order $m$ is a function of the form
\[
s(t) = e^{2\pi j y(t)},
\]
where $j = \sqrt{-1}$, and $t$ is a real number, often representing time, and 
\[
y(t) = \tilde{\mu}_0 +\tilde{\mu}_1 t + \tilde{\mu}_2 t^2 + \dots \tilde{\mu}_m t^m
\]
is a polynomial of order $m$.  In practice the signal is typically sampled at discrete points in `time', $t$. In this paper we only consider uniform sampling, where the gap between consecutive samples is constant. In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal is
\[
s_n = s(n) = e^{2\pi j y(n)},
\] 
where $n$ is an integer.  Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of observations of the noisy sampled signal
\begin{equation}\label{eq:Y_nsamplednoisey}
Y_n = \rho s_n + X_n,
\end{equation}
where $\rho$ is a real number greater than zero representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence complex noise variables. In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}.  It was shown in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.  We discuss this in Section \ref{sec:ident_aliasing}.

@@

An obvious estimator of the unknown coefficients is the least squares estimator (LSE).  This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.  When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the LSE is an effective approach, being both computationally efficient and statistically accurate \cite{Quinn2009_dasp_phase_only_information_loss,Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}\cite[Sec.~6.4~and~9.1]{McKilliam2010thesis}. When $m \geq 2$ the computational complexity of the LSE is large~\cite[Sec.~10.1]{McKilliam2010thesis}\cite{Abatzoglou_ml_chirp_1986}. For this reason many authors have considered alternative approaches to polynomial phase estimation. These can loosely be grouped into two classes, estimators based on polynomial phase transforms, such as the discrete polynomial phase transform (DPT)~\cite{Peleg_DPT_1995,Peleg1991_est_class_PPS_1991,Porat_asympt_HAF_DPT_1996} and the high order phase function~\cite{Farquharson_another_poly_est_2005}, and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, the estimator of Djuric and Kay~\cite{Djuric_phase_unwrap_chirp_1990}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.

In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.  We call this the \emph{least squares unwrapping estimator} (LSU)~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009}\cite[Chap. 8]{McKilliam2010thesis}.  It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance. %Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.  A drawback of the LSU estimator is that computing a nearest lattice point is, in general, computationally difficult.  In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder} \cite{Pohst_sphere_decoder_1981}, and \emph{Babai's nearest plane algorithm} \cite{Babai1986}. The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for $N<50$ \cite{Jalden2005_sphere_decoding_complexity}.  Computing the Babai point requires only $O(N^2)$ operations, but its statistical performance is comparatively poor at low signal to noise ratio (SNR). A major point of interest is that the lattices considered are not \emph{random}, and therefore may admit fast nearest point algorithms. 
%This has been studied in~\cite[Sec.~4.2]{McKilliam2010thesis} where polynomial time algorithms where found that compute the nearest point exactly.  Unfortantely, although polynomial time, the algorithm are still computationally very slow in practice. 
% Two main questions were raised in \cite{McKilliam2009asilomar_polyest_lattice}:
% \begin{enumerate}
% \vspace{-0.1cm}
% \item What are the asymptotic statistical properties of the LSU estimator?
% \vspace{-0.18cm}
% \item Can the structure of the lattice be exploited to obtain fast (exact or approximate) nearest point algorithms and thereby a computationally efficient estimator?
% \vspace{-0.1cm}
% \end{enumerate}
% The purpose of this paper is to answer the first question. We prove that the LSU estimator is strongly consistent and asymptotically normally distributed. 
%We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
In this paper we derive the asymptotic properties of the LSU estimator, showing that it is strongly consistent and asymptotically normally distributed.  Similar results were stated without a complete proof in~\cite{McKilliam_polyphase_est_icassp_2011}.  Here, we give a proof.  The results here are also more general than in~\cite{McKilliam_polyphase_est_icassp_2011}, allowing for a wider class of noise distributions.

The paper is organised in the following way. Section~\ref{sec:lattice-theory} describes the required concepts from lattice theory.  In Section~\ref{sec:ident_aliasing} we briefly describe the identifiable region that was also derived in~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase}.  These identifiability results are required in order to properly understand the statistical properties of polynomial phase estimators. In Section \ref{sec:least-squar-unwr} we describe the LSU estimator and derive its asymptotic statistical properties.
%and also consider techniques for computing the LSU as a nearest lattice point problem. 
Section \ref{sec:simulations} describes the results of Monte Carlo simulations that compare the performance of the LSU estimator with existing estimators.  These simulations agree with the derived asymptotic properties.

@@

A \term{lattice},  $\Lambda$, is a discrete subset of points in $\reals^n$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^d \}
\]
where $\Bbf \in \reals^{n \times d}$ is an $n \times d$ matrix of rank $d$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.  If $n = d$ the lattice is said to be \emph{full rank}.  Lattices are discrete Abelian groups under vector addition.  They are subgroups of the Euclidean group $\reals^n$.  Lattices naturally give rise to \emph{tessellations} of $\reals^n$ by the specification of a set of coset representatives for the quotient $\reals^n / \Lambda$.  One choice for a set of coset representatives is a \emph{fundamental parallelepiped}, i.e., the parallelepiped generated by the columns of a generator matrix.  Another choice is based on the \emph{Voronoi cell}, i.e., those points from $\reals^n$ nearest (with respect to a given norm) to the lattice point at the origin.  It is always possible to construct a \emph{rectangular} set of representatives, as the next proposition will show.  We will use these rectangular regions for describing the aliasing of polynomial phase signals in Section~\ref{sec:ident_aliasing}.  These rectangular regions will also be important for the derivation of the asymptotic properties of the LSU estimator in Section~\ref{sec:least-squar-unwr}.

@@

Given a point $\ybf\in\reals^n$ and a lattice $\Lambda$ contained in~$\reals^n$, the \emph{nearest lattice point problem} is to find the point $\xbf \in \Lambda$ such that the distance (with respect to a given norm) between $\ybf$ and $\xbf$ is minimised.  Here, the Euclidean norm is assumed, so that the nearest point is
\[ 
\arg\min_{\xbf \in \Lambda} \| \xbf - \ybf \|^2.
\]
The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute a nearest lattice point in reasonable time if the dimension is small. One such algorithm, introduced by Pohst~\cite{Pohst_sphere_decoder_1981}, was popularised in the signal processing and communications fields by Viterbo and Boutros~\cite{Viterbo_sphere_decoder_1999} and has since been called the \emph{sphere decoder}.

Approximate algorithms for computing the nearest point have also been studied.  A classic example is \emph{Babai's nearest plane algorithm}~\cite{Babai1986}, which requires $O(n^4)$ arithmetic operations in the worst case where $n$ is the dimension of the lattice and only $O(n^2)$ if the lattice basis is \emph{Lov\'asz reduced}~\cite{Lenstra1982}\index{Lov\'asz reduced}. Recently, various approximate techniques for solving the nearest lattice point problem have been motivated by applications to multiple-input multiple-output communications.  An example is the \emph{$K$-best algorithm}~\cite{Zhan2006_K_best_sphere_decoder} that works similarly to the sequential $M$-algorithm~\cite{Anderson1984_seq_coding_alg} used in coding theory. Yet another example is the \emph{fixed sphere decoder}~\cite{Jalden2009_error_prob_fixed_sphere_decoder,Barbero2008_fixed_sphere_decoder}.

In~\cite{McKilliam2009asilomar_polyest_lattice}, it was shown how the least squares phase unwrapping estimator of a polynomial phase signal can be computed by finding a nearest point in a lattice.  We will briefly describe this result in Section~\ref{sec:least-squar-unwr}.

@@

\section{The Least squares unwrapping estimator}\label{sec:least-squar-unwr}

@@

Figure~\ref{plot:polyest} displays the performance of the LSU estimator, the DPT and Kitchen's estimator when $m=3$. Both the DPT and Kitchen's estimator perform poorly because they do not work correctly for some parameters in the identifiable region $B$. This behaviour has been observed previously~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase,McKilliam_polyphase_est_icassp_2011}. Again the performance of the LSU is well predicted by Theorem~\ref{thm:asymp_proof} provided that the noise variance is small enough to avoid the threshold effect and provided that $N$ is sufficiently large.  The asymptotic CRB is also plotted.

\begin{figure}[tp] 
   	\centering 
  		\includegraphics[width=\linewidth]{code/lseplot-1.mps} 
               %\includegraphics{code/lseplot-1.mps}
   		\caption{Performance of the LSE and LSU estimator for a polynomial phase signal of order two, i.e. $m=2$.  (Top) Mean square error (MSE) in the second order parameter $\mu_2$ for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the $X_1, \dots, X_N$. The $X_1, \dots, X_N$ are independent and identically distributed (i.i.d.) complex Gaussian random variables.  (Bottom) MSE in second order parameter $\mu_2$ versus the variance $\sigma^2$ of the $\Phi_1, \dots, \Phi_N$. The $\Phi_1, \dots, \Phi_N$ are i.i.d. uniform random variables.} 
   		\label{plot:lsetest} 
  \end{figure}

\begin{figure}[p] 
   	\centering 
  		\includegraphics[width=\linewidth]{code/gaussianplot4-1.mps}
                %\includegraphics{code/gaussianplot4-1.mps}
  		\caption{Mean square error (MSE) of polynomial phase estimators for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the complex noise terms $X_1, \dots X_N$. The polynomial phase signal has order 3, and the $X_1, \dots, X_N$ are independent and identically distributed complex Gaussian random variables. (Top left) MSE in the phase coefficient $\mu_0$.  (Top right) MSE in the frequency coefficient $\mu_1$.  (Bottom left) MSE in the quadratic coefficient $\mu_2$.  (Bottom right) MSE in the cubic coefficient $\mu_3$.} 
  		\label{plot:polyest} 
 \end{figure}

@@

since $\expect \expect_\Phi$ is equivalent to $\expect$.   Let $\sigma_n = \frac{g_n}{\abs{g_n}}$ and put $\sigma_n = 1$ if $g_n = 0$.  The $\sigma_n$ are thus independent with $\prob\{ \sigma_n = -1 \} = \tfrac{1}{2}$ and $\prob\{ \sigma_n = 1 \} = \tfrac{1}{2}$.  As $\Phi_n$ and $\Phi_n^\prime$ are independent and identically distributed, the random variable $f_{nN} - f_{nN}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_n( f_{nN} - f_{nN}^\prime)$.  Thus,

>>>>

Polynomial phase signals arise in numerous fields including radar, sonar, geophysics, speech analysis, biology, and radio communication~\cite{Hlawatsch_lin_quad_time_freq_spmag_1992,Ridleyspeechpolyphase1989, Suga_1975_bats_echolocation, Moss_2005echolocation}. In radar and sonar applications polynomial phase signals arise when acquiring radial velocity and acceleration (and higher order motion descriptors) of a target from a reflected signal, and also in continuous wave radar and low probability of intercept radar.  In biology, polynomial phase signals are used to describe the sounds emitted by bats and dolphins for echo location.

A polynomial phase signal of order $m$ is a function of the form
\[
s(t) = e^{2\pi j y(t)},
\]
where $j = \sqrt{-1}$, and $t$ is a real number, often representing time, and 
\[
y(t) = \tilde{\mu}_0 +\tilde{\mu}_1 t + \tilde{\mu}_2 t^2 + \dots \tilde{\mu}_m t^m
\]
is a polynomial of order $m$.  In practice the signal is typically sampled at discrete points in `time', $t$. In this paper we only consider uniform sampling, where the gap between consecutive samples is constant. In this case we can always consider the samples to be taken at some set of consecutive integers and our sampled polynomial phase signal is $s_n = s(n) = e^{2\pi j y(n)}$, where $n$ is an integer.  Of practical importance is the estimation of the coefficients $\tilde{\mu}_0, \dots, \tilde{\mu}_m$ from a number, say $N$, of consequetive observations of the noisy sampled signal
\begin{equation}\label{eq:Y_nsamplednoisey}
Y_n = \rho s_n + X_n,
\end{equation}
where $\rho$ is a real number greater than zero representing the (usually unknown) signal amplitude and $\{X_n, n \in \ints\}$ is a sequence complex noise variables. In order to ensure identifiability it is necessary to restrict the $m+1$ coefficients to a region of $m+1$ dimensional Euclidean space $\reals^{m+1}$ called an \emph{identifiable region}.  It was shown in \cite{McKilliam2009IndentifiabliltyAliasingPolyphase} that an identifiable region tessellates a particular $m+1$ dimensional lattice.  We discuss this in Section \ref{sec:ident_aliasing}.

@@

An obvious estimator of the unknown coefficients is the least squares estimator.  
%This is also the maximum likelihood estimator (MLE) when the noise sequence $\{X_n\}$ is white and Gaussian.  
When $m=0$ (phase estimation) or $m=1$ (frequency estimation) the least squares estimator is an effective approach, being both computationally efficient and statistically accurate~\cite{Hannan1973,Quinn2001,McKilliam_mean_dir_est_sq_arc_length2010}. When $m \geq 2$ the computational complexity of the least squares estimator is large, and for this reason authors have considered alternative estimators. These can loosely be grouped into two classes; estimators based on polynomial phase transforms, such as the discrete polynomial phase transform~\cite{Peleg_DPT_1995,Porat_asympt_HAF_DPT_1996} and the high order phase function~\cite{Farquharson_another_poly_est_2005}; and estimators based on phase unwrapping, such as Kitchen's unwrapping estimator~\cite{Kitchen_polyphase_unwrapping_1994}, and Morelande's Bayesian unwrapping estimator~\cite{Morelande_bayes_unwrapping_2009_tsp}.

In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.  We call this the \emph{least squares unwrapping estimator} (LSU).  It was shown in \cite{McKilliam2009asilomar_polyest_lattice} that the LSU estimator can be represented as a \emph{nearest lattice point problem}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance. %Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.  A drawback of the LSU estimator is that computing a nearest lattice point is, in general, computationally difficult.  In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder} \cite{Pohst_sphere_decoder_1981}, and \emph{Babai's nearest plane algorithm} \cite{Babai1986}. The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for $N<50$ \cite{Jalden2005_sphere_decoding_complexity}.  Computing the Babai point requires only $O(N^2)$ operations, but its statistical performance is comparatively poor at low signal to noise ratio (SNR). A major point of interest is that the lattices considered are not \emph{random}, and therefore may admit fast nearest point algorithms. 
%This has been studied in~\cite[Sec.~4.2]{McKilliam2010thesis} where polynomial time algorithms where found that compute the nearest point exactly.  Unfortantely, although polynomial time, the algorithm are still computationally very slow in practice. 
% Two main questions were raised in \cite{McKilliam2009asilomar_polyest_lattice}:
% \begin{enumerate}
% \vspace{-0.1cm}
% \item What are the asymptotic statistical properties of the LSU estimator?
% \vspace{-0.18cm}
% \item Can the structure of the lattice be exploited to obtain fast (exact or approximate) nearest point algorithms and thereby a computationally efficient estimator?
% \vspace{-0.1cm}
% \end{enumerate}
% The purpose of this paper is to answer the first question. We prove that the LSU estimator is strongly consistent and asymptotically normally distributed. 
%We also apply another known approximate nearest point algorithm, called the $K$-best method \cite{Zhan2006_K_best_sphere_decoder}, that we show provides near sphere decoder performance, but can be computed in a reasonable amount of time if $N$ is less than about 300.
In this paper we derive the asymptotic properties of the LSU estimator, showing that it is strongly consistent and asymptotically normally distributed.  Similar results were stated without proof in~\cite{McKilliam_polyphase_est_icassp_2011}.  Here, we give a proof.  The results here are also more general than in~\cite{McKilliam_polyphase_est_icassp_2011}, allowing for a wider class of noise distributions.

An interesting property is that the estimator of the $k$th polynomial phase coefficient converges almost surely to $\tilde{\mu}_k$ at rate $N^{-k}$.  This is perhaps not surprising, since it is the same rate observed in polynomial regression.  However, asserting that convergence at this rate occurs is not trivial.  For this purpose we make use of elementary results about the number of arithmetic progressions that fit inside subsets of $\{1,2,\dots,N\}$.  This proof technique appears to be novel, and may be useful in other problems.  The proof of asymptotic normality is complicated by the fact that the objective function corresponding with the LSU estimator is not differentiable everywhere.  Techniques developed by Huber~\cite{Huberrobustest1964,Huberbehavmaxnonstandard1967} and refined by Pollard~\cite{Pollard_new_ways_clts_1986,Pollard_asymp_empi_proc_1989} become useful here.  We believe the LSU estimator to be an interesting new application of these techniques.  We are hopeful that the proof techniques developed here will be useful for purposes other than polynomial phase estimation, and in particularly other applications involving data that is `wrapped' in some sense.  Potential candidates are the phase wrapped images observed in synthetic aperture radar and magnetic resonance imaging~\cite{Nico_phaseunwrappingSAR_2000}.

The paper is organised in the following way. Section~\ref{sec:lattice-theory} describes some preliminary concepts from lattice theory, and in Section~\ref{sec:ident_aliasing} we use these results to describe an identifiable region for the set of polynomial phase coefficients.  These identifiability results are required in order to properly understand the statistical properties of polynomial phase estimators. In Section \ref{sec:least-squar-unwr} we describe the LSU estimator and state its asymptotic statistical properties.  Section~\ref{sec:strongconstproof} gives the proof of strong consistency and Section~\ref{sec:centlimitproof} gives the proof of asymptotic normality. 
%and also consider techniques for computing the LSU as a nearest lattice point problem. 
Section~\ref{sec:simulations} describes the results of Monte Carlo simulations with the LSU estimator.  These simulations agree with the derived asymptotic properties.

@@

A \term{lattice},  $\Lambda$, is a discrete subset of points in $\reals^n$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^d \}
\]
where $\Bbf \in \reals^{n \times d}$ is an $n \times d$ matrix of rank $d$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.  If $n = d$ the lattice is said to be \emph{full rank}.  Lattices are discrete Abelian groups under vector addition.  They are subgroups of the Euclidean group $\reals^n$.  Lattices naturally give rise to \emph{tessellations} of $\reals^n$ by the specification of a set of coset representatives for the quotient $\reals^n / \Lambda$.  One choice for a set of coset representatives is a \emph{fundamental parallelepiped}, i.e., the parallelepiped generated by the columns of a generator matrix.  Another choice is based on the \emph{Voronoi cell}, i.e., those points from $\reals^n$ nearest (with respect to a given norm) to the lattice point at the origin.  In what follows when we refer to the Voronoi cell, we will also mean the Vornoi cell with respect to the Euclidean norm.  It is always possible to construct a \emph{rectangular} set of representatives, as the next proposition will show.  We will use these rectangular regions for describing the aliasing of polynomial phase signals in Section~\ref{sec:ident_aliasing}.  These rectangular regions will be important for the derivation of the asymptotic properties of the LSU estimator in Section~\ref{sec:least-squar-unwr}.

@@

% Given a point $\ybf\in\reals^n$ and a lattice $\Lambda$ contained in~$\reals^n$, the \emph{nearest lattice point problem} is to find the point $\xbf \in \Lambda$ such that the distance (with respect to a given norm) between $\ybf$ and $\xbf$ is minimised.  Here, the Euclidean norm is assumed, so that the nearest point is
% \[ 
% \arg\min_{\xbf \in \Lambda} \| \xbf - \ybf \|^2.
% \]
% The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute a nearest lattice point in reasonable time if the dimension is small. One such algorithm, introduced by Pohst~\cite{Pohst_sphere_decoder_1981}, was popularised in the signal processing and communications fields by Viterbo and Boutros~\cite{Viterbo_sphere_decoder_1999} and has since been called the \emph{sphere decoder}.

% Approximate algorithms for computing the nearest point have also been studied.  A classic example is \emph{Babai's nearest plane algorithm}~\cite{Babai1986}, which requires $O(n^4)$ arithmetic operations in the worst case where $n$ is the dimension of the lattice and only $O(n^2)$ if the lattice basis is \emph{Lov\'asz reduced}~\cite{Lenstra1982}\index{Lov\'asz reduced}. Recently, various approximate techniques for solving the nearest lattice point problem have been motivated by applications to multiple-input multiple-output communications.  An example is the \emph{$K$-best algorithm}~\cite{Zhan2006_K_best_sphere_decoder} that works similarly to the sequential $M$-algorithm~\cite{Anderson1984_seq_coding_alg} used in coding theory. Yet another example is the \emph{fixed sphere decoder}~\cite{Jalden2009_error_prob_fixed_sphere_decoder,Barbero2008_fixed_sphere_decoder}.

% In~\cite{McKilliam2009asilomar_polyest_lattice}, it was shown how the least squares phase unwrapping estimator of a polynomial phase signal can be computed by finding a nearest point in a lattice.  We will briefly describe this result in Section~\ref{sec:least-squar-unwr}.

@@

\section{The least squares unwrapping estimator}\label{sec:least-squar-unwr}

@@

Figures~\ref{plot:polyest1}~and~~\ref{plot:polyest2} displays the performance of the LSU estimator, the DPT and Kitchen's estimator when $m=3$. Both the DPT and Kitchen's estimator perform poorly because they do not work correctly for some parameters in the identifiable region $B$. This behaviour has been observed previously~\cite{McKilliam2009IndentifiabliltyAliasingPolyphase,McKilliam_polyphase_est_icassp_2011}. Again the performance of the LSU is well predicted by Theorem~\ref{thm:asymp_proof} provided that the noise variance is small enough to avoid the threshold effect and provided that $N$ is sufficiently large.  The asymptotic CRB is also plotted.

\begin{figure}[p] 
   	\centering 
  		\includegraphics[width=\linewidth]{code/lseplot-1.mps} 
               %\includegraphics{code/lseplot-1.mps}
   		\caption{Performance of the LSE and LSU estimator for a polynomial phase signal of order two, i.e. $m=2$.  (Top) Mean square error (MSE) in the second order parameter $\mu_2$ for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the $X_1, \dots, X_N$. The $X_1, \dots, X_N$ are independent and identically distributed (i.i.d.) complex Gaussian random variables.  (Bottom) MSE in second order parameter $\mu_2$ versus the variance $\sigma^2$ of the $\Phi_1, \dots, \Phi_N$. The $\Phi_1, \dots, \Phi_N$ are i.i.d. uniform random variables.} 
   		\label{plot:lsetest} 
  \end{figure}

\begin{figure}[p] 
   	\centering 
  		\includegraphics[width=\linewidth]{code/gaussianplot4-1.mps}
                %\includegraphics{code/gaussianplot4-1.mps}
  		\caption{Mean square error (MSE) of polynomial phase estimators for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the complex noise terms $X_1, \dots X_N$. The polynomial phase signal has order 3, and the $X_1, \dots, X_N$ are independent and identically distributed complex Gaussian random variables. (Top left) MSE in the phase coefficient $\mu_0$.  (Top right) MSE in the frequency coefficient $\mu_1$.  (Bottom left) MSE in the quadratic coefficient $\mu_2$.  (Bottom right) MSE in the cubic coefficient $\mu_3$.} 
  		\label{plot:polyest1} 
 \end{figure}

\begin{figure}[p] 
   	\centering 
  		\includegraphics[width=\linewidth]{code/gaussianplot4-2.mps}
                %\includegraphics{code/gaussianplot4-1.mps}
  		\caption{Mean square error (MSE) of polynomial phase estimators for $N=10,50$ and $200$ versus the variance $\sigma_c^2$ of the complex noise terms $X_1, \dots X_N$. The polynomial phase signal has order 3, and the $X_1, \dots, X_N$ are independent and identically distributed complex Gaussian random variables. (Top left) MSE in the phase coefficient $\mu_0$.  (Top right) MSE in the frequency coefficient $\mu_1$.  (Bottom left) MSE in the quadratic coefficient $\mu_2$.  (Bottom right) MSE in the cubic coefficient $\mu_3$.} 
  		\label{plot:polyest2} 
 \end{figure}

@@

since $\expect \expect_\Phi$ is equivalent to $\expect$.   Let $\sigma_n = \frac{g_n}{\abs{g_n}}$ and put $\sigma_n = 1$ if $g_n = 0$.  The $\sigma_n$ are independent with $\prob\{ \sigma_n = -1 \} = \tfrac{1}{2}$ and $\prob\{ \sigma_n = 1 \} = \tfrac{1}{2}$.  As $\Phi_n$ and $\Phi_n^\prime$ are independent and identically distributed, the random variable $f_{nN} - f_{nN}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_n( f_{nN} - f_{nN}^\prime)$.  Thus,
